\LARGE \bf Energy-Aware Dynamic Planning Algorithm for Autonomous UAVs

### 2020-12-09 icra-2021.tex
added illustrative abstract
changes in the intro

\usepackage[inline]{enumitem}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{tikz}
\usetikzlibrary{automata,positioning,decorations.pathreplacing}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}

@@

Many scenarios involving unmanned aerial vehicles (UAVs), such as precision agriculture, search and rescue, and surveillance, require high autonomy but have limited energy budgets. A typical example of these scenarios is a UAV flying a path and performing some on-board computational tasks. For instance, the UAV might detect ground patterns and inform other ground-based actors with little human interaction. We refer to the computational tasks that can be dynamically replanned as computations. We are interested in energy-aware dynamic planning, which would find optimal tradeoffs between the path, computations, and energy requirements. Current generic planning solutions for outdoor UAVs do not plan path and computations dynamically, nor are energy-aware. They are often semi-autonomous; the plan--the path and computations--is static and usually defined using planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV planning problem.

Planning algorithms literature for mobile robots includes topics such as trajectory generation and path planning. Generally, the algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}. However, they apply to a small number of robots~\cite{kim2005energy}, and focus exclusively on planning the trajectory for these robots~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being significantly influenced by computations~\cite{mei2005case}. Given the availability of powerful GPU-equipped mobile hardware~\cite{rizvi2017general}, the use of computations is expected to increase in the near future~\cite{abramov2012real,satria2016real,jaramillo2019visual}. More complex planning, which includes a broader concept of the plan being a set of tasks and a path, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}.

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to a planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding,de2017guidance}. We denote the trajectory with a mathematical function that we refer to as the trajectory explicit equation (TEE). The path is composed of multiple TEEs. We express the computations with a modeling tool presented in our previous work~\cite{seewald2019coarse}. The algorithm alters the TEE and computations within some parameters using output model predictive control (MPC). These parameters are the maximum, minimum, and current levels of the TEE and computations. The user specifies the parameters in an initial plan with an energy budget. The energy budget is the battery capacity and some other battery parameters. In the remainder of the paper, we adopt the following notation. We refer to the current TEE and computations parameters simply as TEE and computations parameters. We refer to both as parameters. We refer to the constraints sets that delimit the maximum and minimum levels of the parameters as TEE set and computations sets, and to both as parameters sets. There can be multiple or none computations parameters per each task, and the same applies to the TEE. There is one TEE set but multiple computations sets, one per each computation parameter. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

To achieve the energy-aware dynamic planning, we further introduce and formally proof a periodic energy model that accounts for the uncertainty (atmospheric interference). We use Fourier analysis to derive the model, and state estimation to address the uncertainty. Periodicity is often present due to the path and computations repetitive patterns~\cite{seewald2020mechanical}. Indeed UAV scenarios often iterate over a set of tasks and trajectories (e.g., monitoring or search and rescue). Given that the plan is periodic, we expect the energy to approximately evolve periodically. 
The observation applies broadly; we show different UAV scenarios with periodic energy signals. Moreover, we briefly investigate such signals in some ground-based mobile robots.

In the spirit of reducing costs and resources, we showcase the algorithm using the problem of dynamic planning for a precision agriculture fixed-wing UAV. Precision agriculture is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The plan is structured as follows. Trajectory-wise, the UAV flies in circles and lines covering a polygon. Computationally-wise, it detects obstacles using a convolutional neural network (CNN) and informs grounded mobile robots employed for harvesting. The algorithm alters the plan; it controls the processing rate and the radius of the circles (affecting the distance between the lines). Figure~\ref{fig:overview} shows an initial slice of such a plan. The UAV first heads to a circular TEE with a given radius, which is later reduced as, e.g., windy weather requires the adjustment of the control input to avoid battery depletion. Computations significantly impact the battery, with a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations (see Section~\ref{sec:experimental}).

@@

The algorithm takes as input a user-specified initial plan that consists of different stages and some initial guesses that are necessary for the position and energy estimation. At each stage, the UAV must follow the TEE within TEE set and do the computations within the computations sets. The algorithm outputs the new position, the instantaneous energy consumption, and the control.

@@

@@

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using the TEE parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:overview}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $c_i^0$ (here we again neglect for simplicity the $\sigma=0$ computations parameters), such that $\varphi_i(\mathbf{p}_k,c_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget).

We derive the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=\begin{bmatrix}\partial\varPhi/\partial x & \partial\varPhi/\partial y\end{bmatrix}^T$, and the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

\begin{figure}[h]
  \centering
  \input{\figpath/path2.tikz}
  \caption{Example of the alteration of a TEE parameter}
  \label{fig:overview}
\end{figure}

@@

\leavevmode\thispagestyle{empty}\newpage
\leavevmode\thispagestyle{empty}\newpage

>>>>

\usepackage[inline]{enumitem}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{tikz}
\usetikzlibrary{automata,positioning,decorations.pathreplacing}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry} 
\usepackage{pifont}

@@

Many scenarios involving unmanned aerial vehicles (UAVs), such as precision agriculture, search and rescue, and surveillance, require high autonomy but have limited energy budgets. A typical example of these scenarios is a UAV flying a path and performing some on-board computational tasks. For instance, the UAV might detect ground patterns and inform other ground-based actors with little human interaction. We refer to such computational tasks that can be dynamically replanned and adapted as \emph{computations}. We are interested in the energy optimization of the path and computations under uncertainty and refer to it as energy-aware dynamic planning. Such planning would find optimal tradeoffs between the path, computations, and energy requirements. Current generic planning solutions for outdoor UAVs do not plan the path and computations dynamically, nor are they energy-aware. They are often semi-autonomous: the plan (the path and computations) is static and usually defined using planning software~\cite{daponte2019review}. Such a state of practice has prompted us to propose an \emph{energy-aware dynamic planning algorithm} for UAVs. The algorithm combines and generalizes some of the past body of knowledge on mobile robot planning problems and addresses the increasing \emph{computational demands} and their relation to energy consumption, path, and autonomy for the UAV planning problem.

Planning algorithms literature for mobile robots includes topics such as trajectory generation and path planning. Generally, the algorithms select an energy-optimized trajectory~\cite{mei2004energy}, e.g., by maximizing the operational time~\cite{wahab2015energy}. However, they apply to a small number of robots~\cite{kim2005energy} and focus exclusively on planning the trajectory~\cite{kim2008minimum}, despite compelling evidence for the energy consumption also being significantly influenced by computations~\cite{mei2005case}. Given the availability of powerful GPU-equipped mobile hardware~\cite{rizvi2017general}, the use of computations is expected to increase in the near future~\cite{abramov2012real,satria2016real,jaramillo2019visual}. More complex planning, which includes a broader concept of the plan being a set of tasks and a path, all focus on the trajectory~\cite{mei2005case,mei2006deployment} and apply to a small number of robots~\cite{sadrpour2013mission,sadrpour2013experimental}. For UAVs specifically, rotorcrafts have gained interest in terms of algorithms for energy-optimized trajectory generation~\cite{morbidi2016minimum,kreciglowa2017energy}.

\begin{figure}[t]
  \centering
  \input{\figpath/il_abstract.tikz}
  \caption{Illustrative abstract. An initial plan is replanned in terms of both the path and computations.}
  \label{fig:il-abs}
\end{figure}

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We present here a brief summary of the algorithm. The user specifies the maximum, minimum, and current levels of some parameters relative to the path and computations in an initial plan with an energy budget. In Figure~\ref{fig:il-abs}, there are two parameters. The radius of the trajectory circle is relative to the trajectory. The quality of image detection is relative to the computations. The algorithm replans the initial plan within the constraints and the energy budget using output model predictive control (MPC). The energy budget is the battery capacity and some other battery parameters. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm. To model the path we use multiple trajectories and denote each of them with a mathematical function. In Figure~\ref{fig:il-abs}, the path contains multiple circles and lines. To model the computations we use a profiling tool presented in previous work~\cite{seewald2019coarse}. To guide the UAV we use a vector field~\cite{de2017guidance} that converges smoothly to the planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding,de2017guidance}.

To achieve the energy-aware dynamic planning, we further introduce and formally proof a periodic energy model that accounts for the uncertainty (atmospheric interference). We use Fourier analysis to derive the model, and state estimation to address the uncertainty. Periodicity is often present due to repetitive patterns in the plan~\cite{seewald2020mechanical}. Indeed, UAV scenarios often iterate over a set of tasks and trajectories (e.g., monitoring or search and rescue). Given that the plan is periodic, we expect the energy consumption to approximately evolve periodically. The observation applies broadly; we show different UAV scenarios that can produce such periodic energy signals (evolutions of the energy consumption). Moreover, we briefly investigate such signals in some ground-based mobile robots.

In the spirit of reducing costs and resources, we showcase the algorithm using the problem of dynamic planning for a precision agriculture fixed-wing UAV. Precision agriculture is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The plan is structured as follows. Trajectory-wise, the UAV flies in circles and lines covering a polygon. Computationally-wise, it detects obstacles using a convolutional neural network (CNN) and informs grounded mobile robots employed for harvesting. The algorithm alters the plan; it controls the processing rate and the radius of the circles (affecting the distance between the lines). Figure~\ref{fig:il-abs} shows an initial slice of such a plan. Computations significantly impact the battery, with a potential extension of up to 13 minutes over an hour by switching from the highest to the lowest level of computations (see Section~\ref{sec:experimental}).

@@

The algorithm takes as input a user-specified initial plan that consists of different stages and some initial guesses that are necessary for the position and energy estimation. At each stage, the UAV must follow the trajectory within the trajectory constraints set and do the computations within the computations constraints sets. There is one trajectory constraints set but multiple computations constraints sets, one per each computation parameter. We refer to these omitting sets in the remainder of the paper. The algorithm outputs the new position, the instantaneous energy consumption, and the control.

@@

\begin{figure}[h]
  \centering
  \input{\figpath/path2.tikz}
  \caption{Example of the alteration of a TEE parameter}
  \label{fig:tee1}
\end{figure}

@@

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using the TEE parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:tee1}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $c_i^0$ (here we again neglect for simplicity the $\sigma=0$ computations parameters), such that $\varphi_i(\mathbf{p}_k,c_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget).

We derive the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=\begin{bmatrix}\partial\varPhi/\partial x & \partial\varPhi/\partial y\end{bmatrix}^T$, and the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:tee1}).

@@

\leavevmode\thispagestyle{empty}\newpage
